State Results

### 2023-07-23 reports/state_figures.Rmd
look into the ratio of cases to deaths pre and post omicron

state_res %>%
  filter(state == "MI" & version %in% c("v7")) %>%
 # filter(state %in% sample(corrected$state,5)) %>%
 # filter(biweek >=6) %>%
  group_by(biweek) %>%
  mutate(xmin = min(date),
         xmax = max(date)) %>%
  ungroup() %>%
   ggplot() +
  geom_ribbon(aes(x = date, 
             ymin = exp_cases_lb,
             ymax = exp_cases_ub,
             fill =vlabel),
           #  fill = "#3E3D3D",
             alpha = .5,
             show.legend=FALSE) +
  # geom_ribbon(aes(x = date,
  #                 fill = 'Covidestim',
  #                 ymin = infections.lo,
  #                 ymax = infections.hi),
  #             alpha = .5) +
  # facet_grid(~vlabel, scales = "free_y",
  #             labeller= labeller(vlabel =as_labeller(TeX, default=label_parsed))) +
  scale_x_date(date_breaks = "2 months", 
               date_labels = "%b %Y") +
  theme_c(legend.position="top") +
  # theme_c(axis.text.x = element_text(size = 16),
  #         axis.title.y = element_text(size=13),
  #         plot.title=element_text(face = "bold",
  #                                 hjust = .5, 
  #                                 size = 24,
  #                                 margin=margin(5,5,15,5,'pt')),
  #         strip.background = element_rect(fill = "#393939"),
  #         strip.text = element_text(color =  "white", size = 16),
  #         strip.text.y = element_text(margin = margin(5,5,5,5,'pt'),
  #                                     size = 22,
  #                                     face="bold"),
  #         strip.text.x =  element_text(margin = margin(5,3,5,3,'pt'),
  #                                      face = "bold",
  #                                      size = 22),
  #         plot.subtitle = ggtext::element_markdown(size = 25,margin = margin(3,5,10,5,'pt'),
  #                                      face = "italic"),
  #         legend.position = "top",
  #         legend.text =element_text(size = 17),
  #         legend.title = element_text(size=18, face="bold")) +
  geom_line(aes(x=week, y=share*m,
                color =variant_category,
                group=variant_category),
            data=variant,
            linewidth=1.3) +
  # scale_fill_manual(values =pal,
  #                   labels = TeX(names(pal)),
  #                    breaks='Covidestim',
  #                   name='')  +
  scale_y_continuous(sec.axis = sec_axis( trans=~./m, name="Variant Proportion"),
                     labels = comma) +
  labs(y = "Estimated Infections",
       title = paste0("Estimated Infections by Version in Michigan"),
       color = "SARS-CoV-2 Variant",
       x="Date") +
  guides(color = guide_legend(override.aes = list(linewidth =3),
                              ncol=2,title.position="top")) +
 # scale_color_brewer(palette="Accent") 
  scale_color_manual(values=varpal) +
  scale_fill_manual(values=pal)

@@

>>>>

state_res %>%
  filter(state == "MI" & version %in% c("v7")) %>%
 # filter(state %in% sample(corrected$state,5)) %>%
 # filter(biweek >=6) %>%
  group_by(biweek) %>%
  mutate(xmin = min(date),
         xmax = max(date)) %>%
  ungroup() %>%
   ggplot() +
  geom_ribbon(aes(x = date, 
             ymin = exp_cases_lb,
             ymax = exp_cases_ub,
             fill =vlabel),
           #  fill = "#3E3D3D",
             alpha = .5,
             show.legend=FALSE) +
  scale_x_date(date_breaks = "2 months", 
               date_labels = "%b %Y") +
  theme_c(legend.position="top") +
  geom_line(aes(x=week, y=share*m,
                color =variant_category,
                group=variant_category),
            data=variant,
            linewidth=1.3) +
  # scale_fill_manual(values =pal,
  #                   labels = TeX(names(pal)),
  #                    breaks='Covidestim',
  #                   name='')  +
  scale_y_continuous(sec.axis = sec_axis( trans=~./m, name="Variant Proportion"),
                     labels = comma) +
  labs(y = "Estimated Infections",
       title = paste0("Estimated Infections by Version in Michigan"),
       color = "SARS-CoV-2 Variant",
       x="Date") +
  guides(color = guide_legend(override.aes = list(linewidth =3),
                              ncol=2,title.position="top")) +
 # scale_color_brewer(palette="Accent") 
  scale_color_manual(values=varpal) +
  scale_fill_manual(values=pal)

@@

# Comparing Overlap

Weaknesses in this approach:

- it doesn't distinguish between a very small interval being covered (i.e. when there are very few cases) and a larger interval being covered (i.e. when there are many cases)
- it doesn't consider the width of the PB interval

```{r, fig.height=7.5}

state_overlaps <- subset %>%
  select(-date) %>%
  distinct() %>%
  mutate(lower_overlap = map2_dbl(infections.lo, exp_cases_lb, ~max(.x,.y)),
         upper_overlap = map2_dbl(infections.hi, exp_cases_ub, ~min(.x,.y)),
         overlap = case_when(
           upper_overlap-lower_overlap > 0 ~   upper_overlap-lower_overlap,
           TRUE ~ 0
         )) %>%
  select(contains('overlap'), name, biweek, 
         population, positive, total, exp_cases_lb,
         exp_cases_ub, contains('infections')) %>%
  mutate(overlap_not_norm = overlap,
         overlap = overlap/population,
         fraction_overlap = overlap_not_norm/(infections.hi-infections.lo))

#############
# BY STATE
#############
state_overlaps %>%
  group_by(name) %>%
  summarize(mean_overlap=mean(fraction_overlap)) %>%
  ggplot(aes(y= fct_reorder(name, mean_overlap),
             x =  mean_overlap)) +
  geom_bar(stat="identity") +
  theme_c() +
  labs(y="",
       x="Mean Overlap",
       title = "Mean Overlap with Covidestim by State",
       subtitle = "Overlap Normalized by Width of Covidestim Interval")

##########################
# BY TWO-WEEK INTERVAL
##########################
state_overlaps %>%
  group_by(biweek) %>%
  summarize(mean_overlap=mean(fraction_overlap)) %>%
  left_join(dates) %>%
  group_by(biweek) %>%
  summarize(date = median(date),
            mean_overlap=unique(mean_overlap)) %>%
  ggplot(aes(x=date,
             y =  mean_overlap)) +
  geom_bar(stat="identity", alpha =.7) +
  geom_vline(xintercept=end_old_model,linetype=2,
             color='darkred',
             linewidth=1.1) +
  theme_c() +
  labs(y="",
       x="Mean Overlap",
       title = "Mean Overlap with Covidestim by Time Interval",
       subtitle = "Overlap Normalized by Width of Covidestim Interval") +
  scale_x_date(date_breaks="2 months", date_labels="%b %Y")

##########################
# BY TWO-WEEK INTERVAL
##########################
# state_overlaps %>%
#   mutate(posrate=positive/total) %>%
#   filter(name %in% sample(unique(state_overlaps$name), 5)) %>%
#   ggplot(aes(x=posrate, y = fraction_overlap)) +
#   geom_point(alpha=.8) +
#   theme_c() +
#   # facet_wrap(~name, scales="free_y") +
#   labs(x = "Positivity Rate", y = "Overlap") +
#   scale_y_continuous(labels=comma)
```

```{r, fig.height=12}

set.seed(123)
state_sample <- sample(unique(state_overlaps$name), 16)

state_overlaps %>%
  mutate(posrate=positive/total) %>%
 # filter(name %in% state_sample) %>%
  ggplot(aes(x=posrate, y = fraction_overlap)) +
  geom_point(alpha=.8) +
  theme_c() +
  facet_wrap(~name, scales="free_y") +
  labs(x = "Positivity Rate", y = "Overlap",
       title = "Comparing Fraction of Interval Covered by Positivity Rate",
        subtitle = "Each Point Corresponds to a State-Biweek")

state_overlaps %>%
  mutate(posrate=positive/total) %>%
#  filter(name %in% state_sample) %>%
  ggplot(aes(x=posrate, y = fraction_overlap)) +
  geom_point(alpha=.8) +
  theme_c() +
 # facet_wrap(~name, scales="free_y") +
  labs(x = "Positivity Rate", y = "Overlap",
       title = "Comparing Fraction of Interval Covered by Positivity Rate",
        subtitle = "Each Point Corresponds to a State-Biweek")

```{r, fig.height=7}

state_overlaps %>%
  mutate(testrate=total/population) %>%
  filter(name %in% state_sample) %>%
  ggplot(aes(x=testrate, y = fraction_overlap)) +
  geom_point(alpha=.8) +
  theme_c() +
  facet_wrap(~name, scales="free") +
  labs( x= "Total Tests / Population Size", 
        y = "Fraction of Covidestim Interval\nOverlapping with the PB interval",
        title = "Comparing Fraction of Interval Covered by Testing Rate",
        subtitle = "Each Point Corresponds to a State-Biweek")

state_overlaps %>%
  mutate(testrate=total/population) %>%
  filter(name %in% state_sample) %>%
  ggplot(aes(x=testrate, y = fraction_overlap)) +
  geom_point(alpha=.8) +
  theme_c() +
  labs( x= "Total Tests / Population Size", 
        y = "Fraction of Covidestim Interval\nOverlapping with the PB interval",
        title = "Comparing Fraction of Interval Covered by Testing Rate",
        subtitle = "Each Point Corresponds to a State-Biweek")

## Relationship Between Agreement and Cumulative Incidence (Pre-Omicron)

```{r, fig.height=11}

state_population_link <- "https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/state/detail/SCPRC-EST2019-18+POP-RES.csv"
state_pop <- read_csv(state_population_link)
statecodes <- read_csv(here("data/demographic/statecodes.csv"))

state_pop <- state_pop %>%
    left_join(statecodes, by = c("NAME" = "state")) %>%
    select(population = POPESTIMATE2019,
           state = code,
           name = NAME) %>%
    filter(!is.na(state))

state_deaths <- tar_read(state_deaths, store=targets_dir)

set.seed(123)
state_sample <- sample(unique(state_deaths$state), 5)

state_deaths %>%
#  filter(state %in% state_sample) %>%
  ggplot(aes(x=date,y=deaths/cases)) +
  geom_line() +
  facet_wrap(~state) +
  theme_c() +
  geom_vline(xintercept=end_old_model, color='darkred', linetype=2)

comp_deaths <- state_deaths %>%
  mutate(omicron=ifelse(date <= mdy("12-02-2021"),
                        "pre_omicron", "omicron")) %>%
  group_by(state, omicron) %>%
  summarize(cumulative_deaths = sum(deaths),
            cumulative_cases= sum(cases)) %>%
  pivot_wider(names_from = omicron, 
              values_from = c(cumulative_deaths, cumulative_cases)) %>%
  mutate(pre_deaths_over_cases = cumulative_deaths_pre_omicron/cumulative_cases_pre_omicron,
         omicron_deaths_over_cases = cumulative_deaths_omicron/cumulative_cases_omicron)

comp_deaths %>%
  select(state,pre_deaths_over_cases,omicron_deaths_over_cases) %>%
  pivot_longer(contains("over_cases")) %>%
  group_by(state) %>%
  mutate(pre = value[which(name=="pre_deaths_over_cases")]) %>%
  mutate(name = case_when(
    grepl("omicron", name) ~ "Omicron",
    grepl("pre", name) ~ "Before Omicron"
  )) %>%
  ggplot(aes(y=fct_reorder(state,pre),  x= value, fill=name)) +
  geom_bar(stat='identity', position='dodge') +
  theme_c(legend.title = element_text(face="bold", size=14)) +
  labs(x = TeX("$\\frac{Total\\,Deaths}{Total\\,Cases}$"),
       y="",
       fill = "Time Period",
       title = "Total Deaths Over Total Cases in the Time Period Before Omicron\n Compared to During Omicron") +
  scale_x_continuous(n.breaks=7)

comp_deaths %>%
  select(state,pre_deaths_over_cases,omicron_deaths_over_cases) %>%
  mutate(pct_change =( omicron_deaths_over_cases -pre_deaths_over_cases ) / pre_deaths_over_cases ) %>% 
  ggplot(aes(y=fct_reorder(state,pct_change),  x= pct_change)) +
  geom_bar(stat='identity', position='dodge') +
  theme_c(legend.title = element_text(face="bold", size=11),
          axis.text.x=element_text(size=14),
          plot.subtitle = element_text(size=16,
                                       face="plain", 
                                       margin=margin(0,0,4,0))) +
  labs(x = TeX("$\\frac{Total\\,Deaths}{Total\\,Cases}$"),
       y="",
       fill = "Time Period",
       title = TeX("Percent Change in $\\frac{Total\\,Deaths}{Total\\,Cases}$ from the Time Period Before Omicron"),
       subtitle="Compared to During Omicron") +
  scale_x_continuous(labels=scales::percent)

comp_deaths %>%
  select(state,pre_deaths_over_cases,omicron_deaths_over_cases) %>%
  mutate(pct_change =( omicron_deaths_over_cases - pre_deaths_over_cases ) / pre_deaths_over_cases ) %>%
  left_join(state_overlaps, by = c('state' = 'name')) %>%
  group_by(state) %>%
  summarize(mean_overlap=mean(fraction_overlap),
            pct_change=unique(pct_change)) %>%
  ggplot(aes(x=pct_change,y=mean_overlap)) +
  geom_point(size=2) +
  theme_c(  plot.subtitle = element_text(size=16,
                                       face="plain", 
                                       margin=margin(0,0,4,0))) +
  scale_x_continuous(labels=scales::percent) +
  labs(x="Percent Change", y = "Mean Overlap\n(By State)", 
        title = TeX("Relationship Between the Percent Change in $\\frac{Total\\,Deaths}{Total\\,Cases}$ from the Time Period Before Omicron"),
       subtitle="Compared to During Omicron versus the Mean Overlap")

comp_deaths %>%
  rename(name=state) %>%
  ungroup() %>%
  left_join(state_pop[,c("name", "population")]) %>%
  mutate(frac_infected = cumulative_cases_pre_omicron/population) %>%
  left_join(state_overlaps[,c("name","fraction_overlap", "biweek")]) %>%
  filter(biweek >= 25) %>%
  group_by(name) %>%
  summarize(mean_overlap=mean(fraction_overlap),
            frac_infected=unique(frac_infected)) %>%
  ggplot(aes(x=frac_infected, y =mean_overlap)) +
  geom_point() +
  theme_c()

covidestim_link <- "https://covidestim.s3.us-east-2.amazonaws.com/latest/state/estimates.csv"

covidestim <- read_csv(covidestim_link)

# join data from each source to include dates before and after 2021-12-02
covidestim <- covidestim %>%
    select(date, contains("infections"), state) %>%
    filter(year(date) > 2020) %>%
    mutate(week = week(date), year = year(date))

covidestim %>%
  group_by(state) %>%
  summarize(infections=sum(infections))%>%
  rename(name=state) %>%
  left_join(state_pop[,c("name", "population")]) %>%
  mutate(frac_infected = infections/population) %>%
  left_join(state_overlaps[,c("name","fraction_overlap", "biweek")]) %>%
  filter(biweek >= 25) %>%
  group_by(name) %>%
  summarize(mean_overlap=mean(fraction_overlap),
            frac_infected=unique(frac_infected)) %>%
  ggplot(aes(x=frac_infected, y =mean_overlap)) +
  geom_point() +
  theme_c()
